\documentclass{article}
\title{CSL105 : Discrete Mathematics\\
\large{Minor Examination}\\
\large{Indian Institute of Technology Ropar}\\
\small{Instructor: Dr. Sudarshan Iyengar}
}
\date{March 2017}
\begin{document}
\maketitle
Total Duration : 2 hours \hfill Total Marks : 80 M\\

\underline{\large{Section I \hspace*{7cm} [5 Marks each]}}
\begin{enumerate} 
\item Give an example of a relation which is [reflexive and symmetric] but not [transitive and antisymmetric].
\item Write the Hasse diagram of the relation $R=\{(a,b)/a|b\}$ where $S={1,2,\dots,10}$. 
\item State Pigeon hole principle and its extended/generalized version.
\item What is the condition for a function to be invertible? Explain with an example. 
\item Show that in a group of 5 people, we cannot always guarantee of a presence of three mutual friends or three mutual non-friends. 
\item A box contains 6 red, 8 green, 10 blue, 12 yellow and 15 white balls. What is the minimum
no. of balls we have to choose randomly from the box to ensure that we get 9 balls of same
color?
\item Which of the following statements is/are TRUE for undirected graphs? P: Number of odd
degree vertices is even. Q: Sum of degrees of all vertices is even. 
\item Let G = (V, E) be a loop free undirected graph. Prove that if G contains no cycle of odd
length than G is bipartite.

\end{enumerate}
\vspace*{3cm}


\underline{\large{Section II \hspace*{7cm} [10 Marks each]}}
\begin{enumerate} 
\item Show that among any n + 1 positive integers not exceeding 2n there must be an integer that divides one of the other integers. 
\item Prove that $n^2-1$ is divisible by 8 whenever $n$ is odd positive integer. 
\item Provide a story proof that $\frac{(n!)}{(3!)^k}$ is an integer, given that $n=3k$. Generalize this result.
\item Enumerate all possible non-isomorphic graphs on 4 vertices.

\end{enumerate}
\end{document}
