\documentclass{article}
\title{Minor Exam CSL105}
\date{March 2017}
\begin{document}
\maketitle

\begin{enumerate}
    \item Prove by Induction that $1+\frac{1}{2}+\frac{1}{3}+\dots=\infty$. 
    \item Show that among any n + 1 positive integers not exceeding 2n there must be an integer that divides one of the other integers. 
    \item Show that $$1.2.3+2.3.4+3.4.5+\dots +n(n+1)(n+2)= \frac{n(n+1)(n+2)(n+3)}{4}$$. 
    \item Prove that $n^2-1$ is divisible by 8 whenever $n$ is odd positive integer. 
    \item You need to choose a password which is at least 6 characters and at most 8 characters in length with an added condition that each character is an uppercase letter or a digit. Also, your password must contain at least one digit. In how many ways can you choose your password? 
    \item Provide a story proof that $\frac{(n!)}{(3!)^k}$ is an integer, given that $n=3k$. Generalize this result. 
    \item Let $S=\{1,2,3,\dots\}$. Consider a relation $R=\{(a,b)/a+b\leq 10\}$. Prove or disprove that $R$ satisfies reflexive, symmetric, antisymmetric and transitive properties. 
    \item Give an example of a relation which is [reflexive and symmetric] but not [transitive and antisymmetric]. 
    \item In the matrix representation of a relation, how does one find if the relation is transitive or not? Prove that your method works. 
    \item Write the Hasse diagram of the relation $R=\{(a,b)/a|b\}$ where $S={1,2,\dots,10}$. 
    \item Every sequence of n2 + 1 distinct real numbers contains a subsequence of length n + 1 that is either strictly increasing or strictly decreasing. 
    \item State Pigeon hole principle and its extended/generalized version. 
    \item Define a bijection with an example. 
    \item State well ordering principle. State and Prove Mathematical Induction.
    \item What is the condition for a function to be invertible? Explain with an example. 
    \item Show that in a group of 5 people, we cannot always guarantee of a presence of three mutual friends or three mutual non-friends. 
    \item Enumerate all possible non-isomorphic graphs on 4 vertices.
    \item In how many ways can 12 different books be distributed among 4 children (all 4 of different age) so that (a) each child gets three books ? (b) The two oldest children get four books each and the two youngest get two books each. 
    \item When $n$ is a positive integer greater than 1, show that $^nC_2+^{(n-1)}C_2$ is a perfect square. 
    \item Assignment 1.2 
    \item Assignment 1.5 
    \item Assignment 1.16 
    \item Assignment 2.1 
    \item Assignment 2.15 
    \item Assignment 2.18 
    \item Assignment 3.1 
    \item Assignment 3.14 
    \item Assignment 3.15 







\end{enumerate}
\end{document}
