\documentclass{article}
\usepackage[english]{babel}
\usepackage{color}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{datetime}
\newcommand\tab[1][1cm]{\hspace*{#1}}
\definecolor{indigo}{RGB}{75,0,130}


%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{enumitem} 
\usepackage{amsfonts} 

\title{CSL105 : Discrete Mathematics\\
\large{Major Examination}\\
\large{Indian Institute of Technology Ropar}\\
\small{Instructor: Dr. Sudarshan Iyengar}
}
\date{April 2017}
\begin{document}
\maketitle
Total Duration : 3 hours \hfill Total Marks : 80 M\\

\underline{\large{Section I \hspace*{10cm} [5 Marks each]}}
\begin{enumerate} 
	\item How many 4 digit decimal numbers are there such that 1 is not in the first place, 2 is not in the second, 3 not in third and 4 not in fourth? Prove.    
    
    \item How many positive integer solutions are there for $x_1+x_2+x_3<10$. Explain.
    
   \item Prove that there exists a $k$ such that 11 divides $2^k -1 $.  
    
    \item Is it possible to give an example of a relation which is not a function. Explain.
    
   \item Given two finite sets $A,B$, we say that $A \times B = B \times A$ iff $\_\_\_\_\_\_\_\_$ ? 
   

\item Is complement of a Tree with more than 4 vertices always connected? Give reason for your answer.

\item How many non-isomorphic induced subgraphs does $K_6$ have?

\item Write a statement that is equivalent to $p\rightarrow q$ and prove it with the help of a truth table.

\item What is the rook polynomial of a $3 \times 3$ chess board? Explain.

\item What is the maximum length of a trail in $K_{2n}$?

    
   
   
   
\end{enumerate}
\vspace*{3cm}


\underline{\large{Section II \hspace*{10cm} [10 Marks each]}}
\begin{enumerate} 
 \item Show that whenever 25 girls and 25 boys are seated around a circular table, there will always be a person, both of whose neighbors are boys.
 
 \item Show that $\sum_{v \in V} deg(v)^2 = \sum_{v \in V}\sum_{u \in N(v)}deg(u)$, where $N(v)$ : \{ set of vertices adjacent to v\}.
 
  
 \item Let $G$ be an undirected graph with subgraphs $G_1,G_2$. If $G=G_1 \cup G_2$ and $G_1 \cap G_2 = K_n$ for some $n \in \mathbb{Z}^+$, then show that :
 \begin{align*}
 P(G,\lambda) = \frac{P(G_1,\lambda)P(G_2,\lambda)}{(\lambda)^n}
 \end{align*}
 
 
 
 %\item Show that $9|n^3 +(n+1)^3+(n+2)^3$ through the principle of mathematical induction.
 \item Show that the number of partitions of a positive integer n
where no summand appears more than twice equals the number
of partitions of n where no summand is divisible by 3.
 
 \item Let $E = (e_{ij})_{mxn}$, $F = (f_{ij})_{mxn}$ be two $m\times n $ (0,1)-matrices. We say that $E$ precedes, or
is less than, $F$, and we write $E \leq F$, if $e_{ij} \leq f_{ij}$, for all $1 \leq i <\leq m$, $1 \leq j \leq n$.\\
For $m,n\in \mathbb{Z}^+$, let $A$ be the set of all $m\times n$ (0,1)-matrices. Prove that the "precedes" relation makes $A$ into a poset.


 
 %\item A pair of dice, one red and other other green, are rolled 6 times. We know that the ordered pairs (1, 1), (1, 5), (2, 4), (3, 6), (4, 2), (4, 4), (5, 1) and (5, 5) did not come up. What is the probability that every value came up on both the red die and the green die?
 

  
 
 \end{enumerate}
\end{document}
\grid
\grid
