\documentclass{article}
\usepackage[english]{babel}
\usepackage{color}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{datetime}
\newcommand\tab[1][1cm]{\hspace*{#1}}
\definecolor{indigo}{RGB}{75,0,130}


%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{enumitem} 
\usepackage{amsfonts} 

\begin{document}
\title{Major exam}
\maketitle
\begin{enumerate}
	%counting
 \item How many 4 digit decimal numbers are there such that 1 is not in the first place, 2 is not in the second, 3 not in third and 4 not in fourth? Prove.
 \item What are the total number of passwords with at least 6 digits and at most 8 digits, with a condition that there must at least be one capital letter and one numeral. Prove your answer.

 \item How many positive integer solutions are there for $x_1+x_2+x_3<10$. Explain.

	
	 %PHP/MI

 \item Prove that there exists a $k$ such that 11 divides $2^k -1 $.  



%Functions and Relations
 \item Is it possible to an example of a relation which is not a function. Explain.
 \item Show that for every bijection function $f$, there exists an inverse.
 \item Given two finite sets $A,B$, we say that $A \times B = B \times A$ iff $\_\_\_\_\_\_\_\_$ ?
 
 \item (Give an exercise problem from the text book which is on the stirling's number of the first kind. Note that it is not second kind).

	 %Graph theory
 \item Is complement of a Tree with more than 4 vertices always connected? Give reason for your answer.
 \item A tree has a Hamilton Path iff $\_\_\_\_\_\_\_\_ $ 
 \item How many non-isomorphic induced subgraphs does $K_6$ have?
 \item Derive the chromatic polynomial of a cycle on 5 vertices? 

	 %Logic
 \item Write a statement that is equivalent to $p\rightarrow q$ and prove it with the help of a truth table.
 \item How do you check the divergence of an infinite sequences using quantifiers? Explain.

	 %PIE
 \item What is the rook polynomial of a 3 X 3 chess board? Explain.
 \item Six married couples are to be seated at a circular table. In how many ways can they arrange themselves so that no wife sits next to her husband.
 \item Assignment question

	 %GF
 \item Assignment question
 \item Assignment question
 \item Assignment question



	 

 
\end{enumerate}
  
 
\end{document}

Counting 2
PHP/MI 2
Functions 2
Relations 2

Graph Theory 3
Logic 3
PIE 3
Generating Functions 3


